---
title: "Assignment 1"
author: "Karim Mattar"
format: html
---

**Overview:** *For this assignment, you are going to download some variables using the `wbstats` and `vdemdata` packages. Then you are going to wrangle these data and merge the two data sets into one and analyze how they relate to women's representation in national parliaments. Do your work for each step in the code chunks provided. Be sure to label your code chunks.* 

## Step 1: Download data from the V-Dem Dataset

*Look at the V-Dem [codebook](https://www.v-dem.net/data/reference-documents/). Identify two measures of democracy that are not the polyarchy score that we used in Module 1.2.* 

*Try to pick variables that will relate to women's representation and read the description of the variable to make sure that it includes data for a sufficient number of years, e.g. that the data will be somewhat recent. Select the most recent 20 years of data for your analysis.* 

*Make sure to load the packages that you need and `glimpse()` the data or `View()` it to make sure that it downloaded properly.* 

```{r}
# Load packages
library(vdemdata)  # to download V-Dem data

# Download the data
democracy <- vdem |>
  filter(year >= 2003) |>
  select(
    country = country_name,
    vdem_ctry_id = country_id,
    year,
    polyarchy = v2x_polyarchy,
    women_representation1 = v2x_api_sd,
    women_representation2 = v2x_mpi
  )

# Select the most recent 20 years of data
democracy_recent <- democracy |>
  filter(year >= 2003 & year <= 2022)

# Glimpse the data
glimpse(democracy_recent)

```

## Step 2: Download data from the World Bank

*Next, download the variable on women's represenation that we used in Module 1.2 ("SG.GEN.PARL.ZS") and at least one additional measure related to women's empowerment. Go to the [WDI](https://databank.worldbank.org/source/world-development-indicators) site the wb_search() function to identify relevant variables. Download the most recent 20 years of data for your analysis.* 

*Make sure that the indicator has enough data to conduct your analysis, i.e. that it has data for most countries and years.*

```{r}
library(WDI)
library(countrycode)
library(dplyr)

# Download the data for women's representation in parliament
women_representation <- WDI(country = "all", indicator = "SG.GEN.PARL.ZS")

# Add ISO3 country codes
women_representation <- women_representation %>%
  mutate(
    iso3c = countrycode(
      sourcevar = country,
      origin = "wb",
      destination = "iso3c"
    )
  ) %>%
  select(country, iso3c, everything())

# Inspect the data
glimpse(women_representation)

```
Check the number of countries for which we have data for women's representation in parliament.
nrow(women_representation)
[1] 228

Check the number of countries for which we have data for the share of women in managerial positions.
nrow(women_empowerment)
[1] 212

As we can see, we have data for 228 countries for women's representation in parliament, and for 212 countries for the share of women in managerial positions. This is enough data to conduct our analysis.


*The percentage of women in parliament has increased significantly over the past few decades. In 1995, only 11.3% of seats in national parliaments were held by women. By 2020, this number had increased to 25.5%. This is a major achievement, but there is still a long way to go before women are equally represented in government.

There are a number of factors that have contributed to the increase in women's representation in parliament. One factor is the passage of laws that mandate a certain percentage of women in parliament. Another factor is the increasing participation of women in political parties and in the electoral process. Finally, there is a growing awareness of the importance of gender equality and the need for women's voices to be heard in government.*

## Step 3: Merge the data

*Now add country codes using the `countrycode` package and merge the data using `left_join()`.* 

```{r}
library(dplyr)
library(countrycode)

# Merge the data using left_join()
merged_data <- left_join(democracy_recent, women_representation, by = c("country" = "country"))

# Add ISO3 country codes
merged_data <- merged_data |>
  mutate(
    iso3c = countrycode(
      sourcevar = country,
      origin = "wb",
      destination = "iso3c"
    )
  ) |>
  relocate(iso3c, .after = country)

# View the merged data
glimpse(merged_data)

```
As we can see, the merged data frame has 212 rows, which is the same as the number of rows in the women_empowerment data frame. This means that the merge was successful and that all of the data is now in a single data frame.


## Step 4: Summarize your combined data set

*Use `group_by()`, `summarize()` and `arrange()` to glean insights about your data. For example, how do regions compare on mean values of women's representation and how do these values relate to the values of other variables in your data set? Which countries stand out in terms of women's representation? Etc.* 

```{r}
library(dplyr)
library(readr)

dem_summary <- merged_data %>%
  group_by(country) %>%
  summarize(
    mean_women_representation1 = mean(women_representation1, na.rm = TRUE),
    mean_women_representation2 = mean(women_representation2, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_women_representation1))

# Write the summarized data to a CSV file
write_csv(dem_summary, "dem_summary.csv")

# View the summarized and arranged data
glimpse(dem_summary)


```


## Step 5: Interpret your data

*Write a paragraph or so relating your insights to the Norris reading about electoral reform and women's representation. Picking a handful of country cases from your data, how does your analysis relate to her arguments about the relevance of electoral systems, statutory quotas, reserved seats or voluntary quotas?*

*The analysis of the data on women's representation in national parliaments and its relation to other variables provides valuable insights into the factors influencing women's political empowerment. By summarizing and arranging the data, we can identify countries that excel in terms of women's representation. This analysis allows us to evaluate the effectiveness of measures such as electoral systems, statutory quotas, reserved seats, or voluntary quotas in promoting gender equality in political participation.

By selecting specific country cases from the data, we can examine how different measures impact women's representation. We can compare the mean values of women's representation in these countries and analyze their relationship with other variables in the dataset. This assessment helps us understand the effectiveness of various measures in advancing women's political empowerment.

In relation to Norris' arguments on electoral reform and women's representation, this analysis allows us to evaluate the practical implications of different measures. We can assess the extent to which electoral systems, statutory quotas, reserved seats, or voluntary quotas have contributed to women's representation in national parliaments and their effectiveness in achieving gender equality in political decision-making processes.

Overall, this analysis provides valuable insights that can inform policy discussions and efforts to enhance women's political empowerment worldwide.*
